Carbonate chemistry at Ny-Ålesund

### 2021-03-19 awipev-CO2.Rmd
Refined impacts of formulations K1 and K2

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )

# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4 = pHinsi(pH = pH_durafet$ph_s_dur, 
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
   )

# Make mean for replicates + SD
pH_seafet_mean <- pH_seafet%>%
  dplyr::select(date, ph_s_sf_temp_insi, ph_s_sf_temp_4)%>% 
  dplyr::group_by(date) %>% 
  dplyr::summarise(ph_s_sf_temp_insi_mean = mean(ph_s_sf_temp_insi, na.rm = TRUE),
                   #ph_s_sf_temp_insi_sd = sd(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_temp_4_mean = mean(ph_s_sf_temp_4, na.rm = TRUE)
                   )

pH_durafet_mean <- pH_durafet%>%
  dplyr::select(date, ph_s_dur_temp_fb)%>% 
  dplyr::group_by(date) %>% 
  dplyr::summarize(ph_s_dur_temp_fb_mean = mean(ph_s_dur_temp_fb, na.rm = TRUE)#,
                   #ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE)
                   )

@@

# durafet pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_dur_temp_fb_mean) %>%
  dplyr::rename(ph_s_dur_temp_fb = ph_s_dur_temp_fb_mean) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")

@@

## Impact of choice of constants  {.tabset .tabset-pills}
### Lueker vs Sulpis using means of monthly means
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
The means used are:
`r knitr::kable(tibble(mean_at, mean_ct, mean_pco2, mean_sal, mean_temp), caption = "Mean discrete values.", format = "simple")`

@@

### Lueker vs Sulpis using all values pCO2 and at_calc
```{r impact of constants whole series, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at_calc
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at_calc, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_sf_temp_4, pco2_fb, at_calc, sal_fb, temp_fb)
# Using Lueker  
carb_l <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize Lueker-derived pH at 4°C
carb_l$pH <- pHinsi(pH = carb_l$pH, 
                         ALK = carb_l$ALK, Tinsi = 4, 
                         Tlab = carb_l$T, Pinsi = 0, 
                         S = carb_l$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# Using Sulpis
carb_s <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize Sulpis-derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, 
                         ALK = carb_s$ALK, Tinsi = 4, 
                         Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# calculate difference Lueker - Sulpis
diff_ls <- as_tibble(carb_l - carb_s) %>% 
  dplyr::rename(delta_ph = pH, delta_ct = DIC, delta_OmegaAragonite = OmegaAragonite) %>%
  dplyr::select(delta_ph, delta_ct, delta_OmegaAragonite)
# calculate descriptive statistics
diff_ls_quant <- tibble(pH = as.numeric(summary(diff_ls$delta_ph)),
                        CT = as.numeric(summary(diff_ls$delta_ct * 1e6)),
                        OmegaAragonite = as.numeric(summary(diff_ls$delta_OmegaAragonite))) %>%
  dplyr::mutate(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), .before = 1)
# prepare tibble to compare calculated pH and spec pH
tmp2 <- tibble(ph_s_sf_temp_4 = tmp$ph_s_sf_temp_4, pH_l = carb_l$pH, pH_s = carb_s$pH) %>% 
  tidyr::drop_na(ph_s_sf_temp_4)
# summary of differences
diff_lueker <- as.numeric(summary(tmp2$ph_s_sf_temp_4 - tmp2$pH_l))
diff_sulpis <- as.numeric(summary(tmp2$ph_s_sf_temp_4 - tmp2$pH_s))
lueker_sulpis_table <- tibble(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), diff_lueker, diff_sulpis)
# make plots
p1 <- ggreg(xdata = tmp2$ph_s_sf_temp_4, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 
            title = "Lueker vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_sf_temp_4, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 
            title = "Sulpis vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis") 
p <- p1 + p2
ggsave(filename = "figures/lueker-sulpis.png", plot = p, width = 20, height = 15, units = "cm")
```
The overall means of pCO2, AT (calculated from the AT-salinity relationship), salinity and temperature in the FerryBox were used to calculate pH from measured pCO2 and AT (from a salinity-AT relationship) and from AT and CT. using the formulations of Lueker et al. (2000) and Sulpis et al. (2020) of K1 and K2. The difference induced by the choice of the carbonic acid constants is large for pH, pCO2 and DIC and negligible for OmegaAragonite.

`r knitr::kable(diff_ls_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Sulpis", format = "simple")`

Here we compare pH measured spectrophotometrically with pH calculated from pCO2 and AT using the sets of K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). All pH values were normalized to a temperature of 4°C. Both formulations overestimate spectrophotometric pH 
`r knitr::kable(lueker_sulpis_table, digits = 3, caption = "Difference between spectrophotometric pH and Lueker and Sulpis", format = "simple")`

>>>>

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4 = pHinsi(pH = pH_durafet$ph_s_dur, 
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
   )
# Make mean for replicates + SD
pH_seafet_mean <- pH_seafet %>%
  dplyr::select(date, ph_s_sf_temp_insi, ph_s_sf_temp_4) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarise(ph_s_sf_temp_insi_mean = mean(ph_s_sf_temp_insi, na.rm = TRUE),
                   #ph_s_sf_temp_insi_sd = sd(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_temp_4_mean = mean(ph_s_sf_temp_4, na.rm = TRUE)
                   )
pH_durafet_mean <- pH_durafet %>%
  dplyr::select(date, ph_s_dur_temp_fb, ph_s_dur_temp_4) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarize(ph_s_dur_temp_fb_mean = mean(ph_s_dur_temp_fb, na.rm = TRUE),
                   #ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE),
                   ph_s_dur_temp_4_mean = mean(ph_s_dur_temp_4, na.rm = TRUE)
                   )

@@

# durafet pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_dur_temp_fb_mean, ph_s_dur_temp_4_mean) %>%
  dplyr::rename(ph_s_dur_temp_fb = ph_s_dur_temp_fb_mean,
                ph_s_dur_temp_4 = ph_s_dur_temp_4_mean) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")

@@

## Impact of choice of constants  {.tabset .tabset-pills}
### Lueker vs Sulpis using means of monthly means
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
The means used are:
`r knitr::kable(tibble(mean_at, mean_ct, mean_pco2, mean_sal, mean_temp), digits = c(1, 1, 0, 1, 1), caption = "Mean discrete values.", format = "simple")`

@@

### Lueker vs Sulpis using FerryBox pCO2 and at_calc
```{r impact of constants on FerryBox data, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at_calc
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at_calc, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_dur_temp_4, pco2_fb, at_calc, sal_fb, temp_fb)
# Using Lueker  
carb_l <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize Lueker-derived pH at 4°C
carb_l$pH <- pHinsi(pH = carb_l$pH, 
                         ALK = carb_l$ALK, Tinsi = 4, 
                         Tlab = carb_l$T, Pinsi = 0, 
                         S = carb_l$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# Using Sulpis
carb_s <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize Sulpis-derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, 
                         ALK = carb_s$ALK, Tinsi = 4, 
                         Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# calculate difference Lueker - Sulpis
diff_ls <- as_tibble(carb_l - carb_s) %>% 
  dplyr::rename(delta_ph = pH, delta_ct = DIC, delta_OmegaAragonite = OmegaAragonite) %>%
  dplyr::select(delta_ph, delta_ct, delta_OmegaAragonite)
# calculate descriptive statistics
diff_ls_quant <- tibble(pH = as.numeric(summary(diff_ls$delta_ph)),
                        CT = as.numeric(summary(diff_ls$delta_ct * 1e6)),
                        OmegaAragonite = as.numeric(summary(diff_ls$delta_OmegaAragonite))) %>%
  dplyr::mutate(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), .before = 1)
# prepare tibble to compare calculated pH and spec pH
tmp2 <- tibble(ph_s_dur_temp_4 = tmp$ph_s_dur_temp_4, pH_l = carb_l$pH, pH_s = carb_s$pH) %>% 
  tidyr::drop_na(ph_s_dur_temp_4)
# summary of differences
diff_lueker <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_l))
diff_sulpis <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_s))
lueker_sulpis_table <- tibble(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), diff_lueker, diff_sulpis)
# make plots
p1 <- ggreg(xdata = tmp2$ph_s_dur_temp_4, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 
            title = "Lueker vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_dur_temp_4, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 
            title = "Sulpis vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis") 
p <- p1 + p2
ggsave(filename = "figures/lueker-sulpis.png", plot = p, width = 20, height = 15, units = "cm")
```
The discrete values of pCO2, AT (calculated from the AT-salinity relationship), salinity and temperature in the FerryBox were used to calculate pH from measured pCO2 and AT (from a salinity-AT relationship) using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). The differences induced by the choice of the carbonic acid constants is:
+ large for pH (0.026 pH units on average; always higher with Sulpis than with Lueker)
+ large for CT (3.7 umol/kg on average; always lower with Sulpis than with Lueker)
+ negligible for OmegaAragonite
`r knitr::kable(diff_ls_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Sulpis using discrete Ferrybox data.", format = "simple")`

Here we compare pH measured spectrophotometrically with pH calculated from pCO2 and AT using the sets of K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). All pH values were normalized to a temperature of 4°C. The absloute differences are large, up to 0.14 pH units. On average, both formulations overestimate spectrophotometric pH, by 0.016 pH units with Lueker et al. (2020) and by 0.042 pH units with Sulpis et al. (2020).
`r knitr::kable(lueker_sulpis_table, digits = 3, caption = "Difference between spectrophotometric pH and pH calculated with pCO2 and AT using the formulations for K1 and K2 of Lueker and Sulpis.", format = "simple")`
